Concurrent Programming with CXL

### 2020-06-07 book/paper.tex
Made it fit better

Another important part of the SBS technique is to keep careful track of the
state.  In the case of the reader/writer lock, the state consists of the
following shared variables:
\begin{itemize}
\item \texttt{nreadersEntered}: the number of readers in the critical section;
\item \texttt{nwritersEntered}: the number of writers in the critical section;
\item \texttt{nreadersWaiting}: the number of readers waiting to enter;
\item \texttt{nwritersWaiting}: the number of writers waiting to enter.
\end{itemize}
Each of the
\texttt{rlock\_acquire}, \texttt{rlock\_release},
\texttt{wlock\_acquire}, and \texttt{wlock\_release} methods must maintain
this state.
To wit, check out \texttt{rlock\_acquire}.  It first checks to see if there
are any writers.  If so, it increments \texttt{nreadersWaiting}.  If not,
it increments \texttt{nreadersEntered}.

@@

Let us return now to \texttt{rlock\_acquire}, in the case that the process
found there is a writer in the critical section, it vacates
\texttt{mutex} and procures \texttt{rsema},
blocking on the semaphore associated
with waiting readers.  When later some other process passes the baton to
it by incrementing \texttt{rsema}), it updates the state by decrementing
\texttt{nreadersWaiting} and incrementing \texttt{nreadersEntered}.  It
then invokes \texttt{V\_one()} one more time.

As an example, consider the case where there is a writer in the critical
section and there are two readers waiting.  Let us see what happens when
the writer calls \texttt{wlock\_release}:
\begin{enumerate}
\item After procuring \texttt{mutex},
the sum of the three semaphores is 0.  The writer then decrements
\texttt{nwritersEntered} and calls \texttt{V\_one()}.
\item \texttt{V\_one()} finds that there are no writers in the critical section
and there are two readers waiting.  It therefore vacates \texttt{rsema}.
The sum of the semaphores is now 1, because \texttt{rsema} is 1.  This
guarantees that only a waiting reader, procuring \texttt{rsema}, can enter
the critical section.
\item When it does, the reader decrements \texttt{nreadersWaiting}
from 2 to 1, and increments \texttt{nReadersEntered} from 0 to 1.
The reader finally calls \texttt{V\_one()}.
\item Again, \texttt{V\_one()} finds that there are no writers and
that there are readers waiting, so again it vacates \texttt{rsema},
passing the baton to the one remaining waiting reader.
\item The remaining reader decrements \texttt{nreadersWaiting} from 1 to 0 and
increments \texttt{nReadersEntered} from 1 to 2.
\item Finally, the remaining reader  calls \texttt{V\_one()}.
\texttt{V\_one()} does not find any processes waiting,
and so it vacates \texttt{mutex}.
\end{enumerate}

@@

>>>>

Another important part of the SBS technique is to keep careful track of the
state.  In the case of the reader/writer lock, the state consists of the
following shared variables:
\begin{itemize}
\item \texttt{r\_entered}: the number of readers in the critical section;
\item \texttt{w\_entered}: the number of writers in the critical section;
\item \texttt{r\_waiting}: the number of readers waiting to enter;
\item \texttt{w\_waiting}: the number of writers waiting to enter.
\end{itemize}
Each of the
\texttt{rlock\_acquire}, \texttt{rlock\_release},
\texttt{wlock\_acquire}, and \texttt{wlock\_release} methods must maintain
this state.
To wit, check out \texttt{rlock\_acquire}.  It first checks to see if there
are any writers.  If so, it increments \texttt{r\_waiting}.  If not,
it increments \texttt{r\_entered}.

@@

Let us return now to \texttt{rlock\_acquire}, in the case that the process
found there is a writer in the critical section, it vacates
\texttt{mutex} and procures \texttt{r\_sema},
blocking on the semaphore associated
with waiting readers.  When later some other process passes the baton to
it by incrementing \texttt{r\_sema}), it updates the state by decrementing
\texttt{r\_waiting} and incrementing \texttt{r\_entered}.  It
then invokes \texttt{V\_one()} one more time.

As an example, consider the case where there is a writer in the critical
section and there are two readers waiting.  Let us see what happens when
the writer calls \texttt{wlock\_release}:
\begin{enumerate}
\item After procuring \texttt{mutex},
the sum of the three semaphores is 0.  The writer then decrements
\texttt{w\_entered} and calls \texttt{V\_one()}.
\item \texttt{V\_one()} finds that there are no writers in the critical section
and there are two readers waiting.  It therefore vacates \texttt{r\_sema}.
The sum of the semaphores is now 1, because \texttt{r\_sema} is 1.  This
guarantees that only a waiting reader, procuring \texttt{r\_sema}, can enter
the critical section.
\item When it does, the reader decrements \texttt{r\_waiting}
from 2 to 1, and increments \texttt{r\_entered} from 0 to 1.
The reader finally calls \texttt{V\_one()}.
\item Again, \texttt{V\_one()} finds that there are no writers and
that there are readers waiting, so again it vacates \texttt{r\_sema},
passing the baton to the one remaining waiting reader.
\item The remaining reader decrements \texttt{r\_waiting} from 1 to 0 and
increments \texttt{r\_entered} from 1 to 2.
\item Finally, the remaining reader  calls \texttt{V\_one()}.
\texttt{V\_one()} does not find any processes waiting,
and so it vacates \texttt{mutex}.
\end{enumerate}

@@

\chapter{Monitors}
